---
title: "Unit 03 Assignment"
author: "Kenneth Richardson"
class: "MSDS 6306"
date: "January 31, 2019"
output: html_document
keep_md: TRUE
---


# 1) GitHub Cloning
    *Using Git, clone the following GitHub repository to your local machine:         https://github.com/caesar0301/awesome-public-datasets.  In RMarkdown, please show the code (commented out, as it’s not R syntax) that you used to create a new directory, navigate to the appropriate directory, and clone the repository to it.  One Git command per line, please

```{r Github, echo=TRUE}
#$ cd 'C:\Users\Kenneth\Documents\SMU\DDS\Unit3'
# git clone https://github.com/caesar0301/awesome-public-datasets.git
```

# 2) Data Summary
  *In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status
  *Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males.  Be sure to give an accurate title and label the axes.
  *Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival.  Make sure the output is a real number for all three means

```{r plot,echo=TRUE}
#a) reading data set
df <- read.csv("~/SMU/DDS/Unit3/awesome-public-datasets/Datasets/titanic.csv/titanic.csv")
head(df)
#b) Plotting frequency of Females and Males
plot(df$Sex, main="Frequency of Females and Males", xlab = "Sex", ylab ="Count")
#c) Mean Age, Fare, Survived
Select.variables <- df[,c("Age","Fare","Survived")]
Select.variables
mean.test <- sapply(Select.variables, mean,na.rm=T )
mean.test
#output     
#Age       Fare   Survived 
#29.6991176 32.2042080  0.3838384 
```

# 3a) Function Building
  *Questions 3A through 3D should be answered in function(x){}.  3E can be outside of the function.
  *Create objects for the median Age, the minimum and maximum Duration of sleep, and the mean and standard deviation of the Rosenberg Self Esteem scale (RSES).  You may need to specify a few options like in Problem 2 and live session

```{r function, echo=TRUE}
# reading sleep data
sleep_data <- read.csv("~/SMU/DDS/Unit3/sleep_data_01.csv")
sleep_data.function <- function(x) {
  Age.median <- median(sleep_data$Age,na.rm = TRUE)
  Dur.min <- min(sleep_data$Duration,na.rm = TRUE)
  Dur.max <- max(sleep_data$Duration,na.rm = TRUE)
  RSES.mean <- mean(sleep_data$RSES,na.rm = TRUE)
  RSES.sd <- sd(sleep_data$RSES,na.rm = TRUE)
  return(
    data.frame(
      Age.median,
      Dur.min,
      Dur.max,
      RSES.mean,
      RSES.sd
    )
  )
}
 #output
sleep_data.function(x)

```

# 3b) Function Building
  *Create a data.frame object called report: it should consist of the median age, the RSES mean and standard deviation respectively divided by five (since there are five questions and these scores are summed), and the range of Duration (the statistical definition of range; it should be a single number.)

```{r Dataframe, echo=TRUE}

report <- function(x){
  Age.median.2 <- sleep_data.function()$Age.median/5
  RSES.mean.2 <- sleep_data.function()$RSES.mean/5
  RSES.sd.2 <- sleep_data.function()$RSES.sd/5
  Range.Dur <- (sleep_data.function()$Dur.max-sleep_data.function()$Dur.min)
  return(
    data.frame(
      Age.median.2,
      RSES.mean.2,
      RSES.sd.2,
      Range.Dur
    )
  )
}
# output
report(x)

```


# 3c) Function Building
  *Change the column names of this data.frame to MedianAge, SelfEsteem, SE_SD, and DurationRange.

```{r Function, echo=TRUE}
change.names <- function(x){
  colnames(x) <- c("MedianAge","SelfEsteem","SE_SD","DurationRange")
  return(x)
}
report2 <- change.names(report(x))
#output
report2

```



# 3d) Function Building
  *Round the report to at most 2 digits: leave this as the closing line to the function.

```{r Function, echo=TRUE}
report.rounded <- function(x){
  round(x,digits = 2)
}
```


# 3e) Function Building
  *Finally, run the function on your sleep data to show the output.

```{r Function, echo=TRUE}
report.rounded(report2)

````

# 4a) FiveThirtyEight Data
  *Install the fivethirtyeight package.

``` {r Install, echo=TRUE]}
#install.packages("fivethirtyeight")
```


# 4b) FiveThirtyEight Data
  *In the listing of Data sets in package ‘fivethirtyeight,’ assign the 22nd data set to an object ‘df.’
```{r Assign, echo=TRUE}
FTE <- data(package = "fivethirtyeight")
df <- FTE$results[22,"Item"]
df
```

# 4c) FiveThirtyEight Data
  *Use a more detailed list of the data sets to write out the URL in a comment to the related news story.
```{r URL, echo=TRUE}
library(fivethirtyeight)
vignette("fivethirtyeight", package = "fivethirtyeight")
#URL Commented
# http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
````

# 4d) FiveThirtyEight Data
  *Using R command(s), give the dimensions and column names of this data frame.


```{r column_dim, echo=TRUE}
fte <- fivethirtyeight::college_recent_grads
# column names
names(fte)
# get output number of Rows, Number of Columns
dim(fte)
```


# 5a) Data Summary
  *Write an R command that gives you the column names of the data frame.  Right after that, write one that counts the number of columns but not rows.  Hint: The number should match one of your numbers in Question 1d for dimensions.

```{r col_name, echo=TRUE}
# column names
names(fte)
# number of columns
ncol(fte)
```


#5b) Data Summary
  *Generate a count of each unique major_category in the data frame.  I recommend using libraries to help.  To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset.  Assign it to major_count.

```{r Count_Majors, echo=TRUE}
  #  dplyr to list unique major and get count of each major
  major_count <- dplyr::count(fte,fte$major_category)
  # renaming columns
  colnames(major_count) = c("Major","Count")
  # output
  major_count
```



#5c) Data Summary



#5d) Data Summary
  *Write the fivethirtyeight data to a csv file.  Make sure that it does not have row labels.
```{r csv_write, echo=TRUE}
require(RCurl)
#  fivethirtyeight Raw data
raw.file <- read.csv(text=getURL("https://raw.githubusercontent.com/rudeboybert/fivethirtyeight/master/data-raw/college-majors/recent-grads.csv"), header=T)
# writing get hub data to a file named Recentgrad
write.csv(raw.file, file = 'C:\\Users\\Kenneth\\Documents\\SMU\\DDS\\unit3\\Recentgrad.csv', row.names = FALSE)
```

# 6)
  *Start a new repository on GitHub for your SMU MSDS homework.  On your local device, make sure there is a directory for Homework at the minimum; you are welcome to add whatever you would like to this repo in addition to your requirements here.
  *Create a README.md file which explains the purpose of the repository, the topics included, the sources for the material you post, and contact information in case of questions. Remember, the one in the root directory should be general.  You are welcome to make short READMEs for each assignment individually in other folders.
  *In one (or more) of the nested directories, post your RMarkdown script, HTML file, and data from ‘fivethirtyeight.’  Make sure that in your README or elsewhere that you credit fivethirtyeight in some way.
  *In your RMarkdown script, please provide the link to this GitHub so the grader can see it.
```{r codebook, echo=TRUE}
# Github link: https://github.com/Kenscieroyal18/SMU-MSDS-Homework
```


